---
title: "Part6-GLMM for endpoint measurement"
author: "LeeLK"
date: "2025-04-08"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Normality and homogeneity test
```{r }
load(file="output/growth_metric_all.RData")

#to check if data distribution is normal
options(digits=2)
metric_all %>% normality () %>% arrange(desc(p_value)) 
metric_all %>% normality () %>% filter(p_value >= 0.05) 

metric_all %>% plot_normality () #to visualize data distribution

find_skewness(metric_all, value=TRUE)#check for skewness value
find_skewness(metric_all, index=FALSE)#check for which variables have heavy skewness

#for leveneTest, use lapply across measurements
#combine light and nutrient level into single factor for grouping
metric_all <- metric_all %>%
  unite("trtmnt", light:nutrient, remove=FALSE)

#name trtmnt as factor
metric_all$trtmnt <- as.factor(metric_all$trtmnt)

#define function to perform Levene's test
levene_test_function <- function(variable) {
  leveneTest(variable ~ trtmnt, data = metric_all)
}

# List of variables to test
variables_to_test <- c("shell_length", "shell_width", "shell_height", 
                       "total_wet_weight_g", "tissue_wet_weight_g",
                       "shell_dry_weight_g", "strength_N", "cell_tissue_g", "chl_a_g")

# Apply the levene_test_function to each variable
results <- lapply(variables_to_test, function(var) {
  levene_test_function(metric_all[[var]])
})

# Name the list elements for easier interpretation
names(results) <- variables_to_test
```

### Data transformation
Data transformation of log10 or sqrt to check if data become normal
```{r}
#log10 transformation
metric_all_trans <- metric_all %>%
  mutate(across(shell_length:chl_a, ~ log1p(.)))

#check normality of log10 transformed data
metric_all_trans %>% normality () %>% filter(p_value >=0.05)
#adj_height, adj_strength, cell_count

#sqrt transformation
metric_all_trans2 <- metric_all %>%
  mutate(across(shell_length:chl_a, ~ sqrt(.)))

#check normality of log10 transformed data
metric_all_trans2 %>% normality () %>% filter(p_value >=0.05)

#log10 transformation
metric_all <- metric_all %>%
  mutate(log_cell_count = log10(cell_count))

#describe the cell count and chl-a
metric_all %>%
  group_by(light, nutrient)%>%
  describe() %>%
  filter(described_variables == "cell_count")
```
## GLMM for end-point measurement
Most variables does not conform to normality even after transformation (except for log10 transformed of cell count), proceed with GLMM which have more flexibility than conventional two-way ANOVA (or its non-parametric version)
```{r}
#load metric_all file
load(file="output/growth_metric_all.RData")

#package
library(tidyverse)
library(rstatix)
library(lme4)
library(sjPlot)
library(gtsummary)
library(emmeans)
library(multcomp)
library(car)

#format variables to factor level
metric_all <- metric_all %>%
  mutate_at(vars(light, nutrient, tank), as.factor)

#run GLMM, with tank as random effect
glmm_1 <- glmer(shell_length ~ light*nutrient + (1|tank), family = Gamma(link ="log"), data =metric_all)

#use sjPLot to check summary/extract the result on random effect
tab_model(glmm_1)

#Anova with wald chi-sq test
#Type II p-values are calculated using the sums of squares for each main effect conditional on the other main effects. 
#Type III p-values conditions the sums of squares on the interaction term as well.

Anova(glmm_1, type=3)

#emmeans for light
emms_1 <- emmeans(glmm_1, pairwise ~ light*nutrient, adjust = "bonferroni")

#alphabet for pairwise
emms1_label <- cld(emms_1, alpha=0.05, Letters=letters, adjust="bonferroni")  

#write csv
write.csv(emms1_label, file ="output/emms1_label.csv")

#run GLMM with adjusted width
glmm_2 <- glmer(adj_width ~ light*nutrient + (1|tank), family = Gamma(link ="log"), data =metric_all)

#use sjPLot to check summary/extract the result on random effect
tab_model(glmm_2)

#Anova with wald chi-sq test
Anova(glmm_2,  type=3)

#emmeans
emms_2 <- emmeans(glmm_2, pairwise ~ light*nutrient, adjust = "bonferroni")

#run GLMM with adj_height
glmm_3 <- glmer(adj_height ~ light*nutrient + (1|tank), family = Gamma(link ="log"), data =metric_all)

#use sjPLot to check summary/extract the result on random effect
tab_model(glmm_3)

#Anova with wald chi-sq test
Anova(glmm_3, type =3)

#emmeans
emms_3<- emmeans(glmm_3, pairwise ~ light*nutrient, adjust = "bonferroni")

#run GLMM with adj_weight
glmm_4 <- glmer(adj_weight ~ light*nutrient + (1|tank), family = Gamma(link ="log"), data =metric_all)

#use sjPLot to check summary/extract the result on random effect
tab_model(glmm_4)

#Anova with wald chi-sq test
Anova(glmm_4, type= 3)

#emmeans
emms_4<- emmeans(glmm_4, pairwise ~ light*nutrient, adjust = "bonferroni")

#alphabet for pairwise
emms4_label <- cld(emms_4, alpha=0.05, Letters=letters, adjust="bonferroni")  

#write csv
write.csv(emms4_label, file ="output/emms4_label.csv")

#run GLMM with total biomass
glmm_5 <- glmer(total_wet_weight_g ~ light*nutrient + (1|tank), family = Gamma(link ="log"), data =metric_all)

#use sjPLot to check summary/extract the result on random effect
tab_model(glmm_5)

#Anova with wald chi-sq test
Anova(glmm_5, type= 3)

#emmeans
emms_5<- emmeans(glmm_5, pairwise ~ light*nutrient, adjust = "bonferroni")

#alphabet for pairwise
emms5_label <- cld(emms_5, alpha=0.05, Letters=letters, adjust="bonferroni")  

#write csv
write.csv(emms5_label, file ="output/emms5_label.csv")

#run GLMM with tissue biomass
glmm_6 <- glmer(tissue_wet_weight_g ~ light*nutrient + (1|tank), family = Gamma(link ="log"), data =metric_all)

#use sjPLot to check summary/extract the result on random effect
tab_model(glmm_6)

#Anova with wald chi-sq test
Anova(glmm_6, type =3)

#emmeans
emms_6<- emmeans(glmm_6, pairwise ~ light*nutrient, adjust = "bonferroni")

#alphabet for pairwise
emms6_label <- cld(emms_6, alpha=0.05, Letters=letters, adjust="bonferroni")  

#write csv
write.csv(emms6_label, file ="output/emms6_label.csv")

#run GLMM with shell strength
glmm_7 <- glmer(adj_strength ~ light*nutrient + (1|tank), family = Gamma(link ="log"), data =metric_all)

#use sjPLot to check summary/extract the result on random effect
tab_model(glmm_7)

#Anova with wald chi-sq test
Anova(glmm_7, type =3)

#emmeans
emms_7<- emmeans(glmm_7, pairwise ~ light*nutrient, adjust = "bonferroni")

#alphabet for pairwise
emms7_label <- cld(emms_7, alpha=0.05, Letters=letters, adjust="bonferroni")  

#write csv
write.csv(emms7_label, file ="output/emms7_label.csv")

#log10 transformation for cell_count
metric_all <- metric_all %>%
  mutate(log_cell_count = log10(cell_count))

#use lmer
lmm_8 <- lmer(log_cell_count ~ light*nutrient + (1|tank), data =metric_all)

#use sjPLot to check summary/extract the result on random effect
tab_model(lmm_8)

#Anova with wald chi-sq test
Anova(lmm_8, type =3)

#emmeans
emms_8<- emmeans(lmm_8, pairwise ~ light*nutrient, adjust = "bonferroni")

#run GLMM with chl_a
glmm_9 <- glmer(chl_a ~ light*nutrient + (1|clam), family = Gamma(link ="log"), data =metric_all)

#use sjPLot to check summary/extract the result on random effect
tab_model(glmm_9)

#Anova with wald chi-sq test
Anova(glmm_9, type=3)

#emmeans
emms_9<- emmeans(glmm_9, pairwise ~ light*nutrient, adjust = "bonferroni")

#alphabet for pairwise
emms9_label <- cld(emms_9, alpha=0.05, Letters=letters, adjust="bonferroni")  

#write csv
write.csv(emms9_label, file ="output/emms9_label.csv")

#run GLMM, with tank as random effect
glmm_ex <- glmer(cond_index ~ light*nutrient + (1|tank), family = Gamma(link ="log"), data =shell_morpho3)

Anova(glmm_ex, type =3)

#emmeans
emms_10<- emmeans(glmm_ex, pairwise ~ light*nutrient, adjust = "bonferroni")
```

## P value triangular matrix
Use triangular matrix to show the p-value summary across variables
```{r}
#extract the p-value from ANOVA result and create tringular matrix
library(reshape2)
library(RColorBrewer)

#
pvalue_matrix <- data.frame(
  variables = c('shell_length', 'shell_adj_width', 'shell_adj_height', 'shell_adj_weight', 
                'shell_adj_strength', 'total_biomass', 'tissue_biomass', 'log_cell_count', 'chl_a'),
  light = c(0.002, 0.72, 0.87, 0.002, 0.1, 0.0001, 0.0001, 0.84, 0.43),
  nutrient = c(0.02, 0.83, 0.3, 0.17, 0.03, 0.04, 0.002, 0.41, 0.03),
  light_nutrient = c(0.12, 0.98, 0.84, 0.1, 0.29, 0.064, 0.08, 0.62, 0.13)
)

# Convert to matrix format
pvalue_matrix_melt <- melt(pvalue_matrix, id.vars = 'variables')

# Create a custom color scale based on p-value thresholds
pvalue_matrix_melt$significance <- cut(pvalue_matrix_melt$value,
                              breaks = c(-Inf, 0.001, 0.01, 0.05, Inf),
                              labels = c("< 0.001", "< 0.01", "< 0.05", "> 0.05"))

#generate plot for pvalue matrix
plot10<-ggplot(pvalue_matrix_melt, aes(x = variable, y = variables, fill = significance)) +
  geom_tile(color = "white",alpha=0.5) +
  scale_fill_manual(values = c("< 0.001" = "red", 
                               "< 0.01" = "orange", 
                               "< 0.05" = "yellow", 
                               "> 0.05" = "gray"), 
                    name = "p-value") +
  geom_text(aes(label = sprintf("%.3f", value)), color = "black", size = 2) +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  coord_fixed() +
  theme(legend.position = "top")
```

## Assemble plot

```{r}
#Fig 4
fig6 <- plot1+plot2+plot3+plot4+plot5+plot6+plot7+plot8+plot9+
  plot_annotation(tag_levels = "a")+
  plot_layout(ncol=3, guides = "collect") & 
  theme(legend.position = "top")

fig6.1 <- wrap_plots(fig6, plot10) +  plot_annotation(tag_levels = "a")
```
