---
title: "Part5-EndPoint_measurement"
author: "LeeLK"
date: "2025-04-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Library
library(tidyverse)
library(readxl)
library(dlookr)#for summary statistics
library(car)#LeveneTest
library(ggsci)
library(ggpubr)
library(broom)
library(patchwork)
library(rstatix)
library(lme4)
library(sjPlot)
library(gtsummary)
library(emmeans)
```
## Initial morphometric
```{r}
#start point measurement
start_morpho <- read_excel("C:/Users/yukin/Desktop/MyWork2025/M6. Tcro_CEA/2. DataAnalysis/Tcro_CEA/analysis_data2.xlsx", sheet = "start_morpho")

#calculate descriptive stats
start_morpho %>% describe ()

#descriptive stats by tank
start_morpho %>% group_by (tank) %>%
  describe ()
```

## Summary for Shell morphometric
```{r}
shell_morpho <- read_excel("C:/Users/yukin/Desktop/MyWork2025/M6. Tcro_CEA/2. DataAnalysis/Tcro_CEA/analysis_data2.xlsx", sheet = "morphometric")

shell_strength <- read_excel("C:/Users/yukin/Desktop/My Work/My Research/Manuscripts/M6. Tcro_CEA/2. DataAnalysis/Tcro_CEA/analysis_data2.xlsx", sheet = "shell_strength")

#join shell_morpho and shell_strength by tank and clam ID
shell_morpho2<- left_join(shell_morpho, shell_strength, by=c("tank","clam"))

save(shell_morpho2, file ="output/shell_morpho2.RData")
```

## Growth (gain)
Only average shell length and total wet weight of start and end were used
```{r}
#select only column 1:4, 6 and 9, from morpho dataset, 6 is end shell length and 9 is end wet weight
#rename the column
initial_end_morpho <- shell_morpho %>%
  select(1:5,9) %>%
  rename(end_length = sh_length, end_weight = total_weight) %>%
  unite("trtmnt", light:nutrient, remove=FALSE)

#cbind 
initial_end_morpho <- cbind(initial_end_morpho, start_morpho)

#remove unwanted column
initial_end_morpho <- initial_end_morpho %>% select(!8)

initial_end_morpho %>% group_by (trtmnt) %>% 
  describe()

#calculate percentage gain for shell length and weight
percent_gain <- initial_end_morpho %>% 
  group_by(trtmnt) %>%
  summarise(across(end_length:start_weight, list(mean = mean), na.rm= TRUE))%>%
  mutate(length_gain = ((end_length_mean - start_length_mean)/start_length_mean)*100) %>%
  mutate(mass_gain = ((end_weight_mean - start_weight_mean)/start_weight_mean)*100)
```

## Adjusted morphometric summary
Add adjusted shell morphometric (adjusted to shell length)
```{r}
#load shell_morpho2 file
load(file="output/shell_morpho2.RData")

#shell variable need to be adjusted with shell length, add adjusted variable using across and a function
#calculate CI, condition index with 
#CI= (Shell Dry Weight/Tissue Wet Weight)*100
shell_morpho3 <- shell_morpho2 %>%
  mutate(across(shell_width:strength_N, ~ .x / shell_length, .names = "adj_{col}")) %>%
  rename(adj_width = adj_shell_width, adj_height = adj_shell_height,
         adj_weight = adj_shell_dry_weight_g, adj_strength=adj_strength_N) %>%
  mutate(cond_index = (tissue_wet_weight_g/shell_dry_weight_g)*100)

#
save(shell_morpho3, file="output/shell_morpho3.RData")
```
## Plotting
```{r}
#load shell morpho3, which have the adjusted variables
load(file="output/shell_morpho3.RData")

#shell length
plot1 <- ggplot(shell_morpho3, aes(x = nutrient, y = shell_length, fill = nutrient)) +
  geom_boxplot()+
  facet_wrap(~ light) +  
  scale_fill_npg(alpha=0.5) +
  labs(y = "ShLength (mm)") +
  theme_pubr() +
  theme(legend.position = "top",
        axis.text.x = element_blank(),
    axis.title.x = element_blank())

#adj shell width
plot2 <- ggplot(shell_morpho3, aes(x = nutrient, y = adj_width , fill = nutrient)) +
  geom_boxplot() +
  facet_wrap(~ light) +  
  scale_fill_npg(alpha = 0.5) +
  labs(y = "ShWidth/Length") +
  theme_pubr() +
  theme(legend.position = "top",
        axis.text.x = element_blank(),
    axis.title.x = element_blank())

#adj shell height
plot3 <- ggplot(shell_morpho3, aes(x = nutrient, y = adj_height, fill = nutrient)) +
  geom_boxplot() +
  facet_wrap(~ light) +  
  scale_fill_npg(alpha = 0.5) +
  labs(y = "ShHeight/Length") +
  theme_pubr() +
  theme(legend.position = "top",
        axis.text.x = element_blank(),
    axis.title.x = element_blank())

#shell dry weight
plot4 <- ggplot(shell_morpho3, aes(x = nutrient, y = adj_weight, fill = nutrient)) +
  geom_boxplot() +
  facet_wrap(~ light) +  
  scale_fill_npg(alpha = 0.5) +
  labs(y = "ShWeight/Length") +
  theme_pubr() +
  theme(legend.position = "top",
        axis.text.x = element_blank(),
    axis.title.x = element_blank())

#adj strength
plot5 <- ggplot(shell_morpho3, aes(x = nutrient, y = adj_strength, fill = nutrient)) +
  geom_boxplot() +
  facet_wrap(~ light) +  
  scale_fill_npg(alpha = 0.5) +
  labs(y = "ShStrength/Length") +
  theme_pubr() +
  theme(legend.position = "top",
        axis.text.x = element_blank(),
    axis.title.x = element_blank())

#shell length
plot_ex <- ggplot(shell_morpho3, aes(x = nutrient, y = cond_index, fill = nutrient)) +
  geom_boxplot()+
  facet_wrap(~ light) +  
  scale_fill_npg(alpha=0.5) +
  labs(y = "Condition Index (mm)") +
  theme_pubr() +
  theme(legend.position = "top",
        axis.text.x = element_blank(),
    axis.title.x = element_blank())
```
## Cell density and chl-a
```{r}
load(file="output/shell_morpho3.RData")

#load chl_a and cell density data
chl_a <- read_excel("C:/Users/yukin/Desktop/My Work/My Research/Manuscripts/M6. Tcro_CEA/2. DataAnalysis/Tcro_CEA/analysis_data2.xlsx", sheet = "chl-a")

cell_density <- read_excel("C:/Users/yukin/Desktop/My Work/My Research/Manuscripts/M6. Tcro_CEA/2. DataAnalysis/Tcro_CEA/analysis_data2.xlsx", sheet = "cell_density")

#join chl_a and cell density to shell_morpho3
metric_all<- shell_morpho3 %>%
  left_join(cell_density, by=c("tank","clam")) %>%
    left_join(chl_a, by=c("tank","clam"))

#save this datafile for next section for normality, ordination 
save(metric_all, file="output/growth_metric_all.RData")

#Morphometric summary
options(digits=3)

metric_all %>%
  select(tank, light, nutrient) %>%
  group_by (light, nutrient) %>% 
  describe () %>%
    select(described_variables, n,light, nutrient, mean, sd)

load(file="output/growth_metric_all.RData")

#total biomass
plot6<- ggplot(metric_all, aes(x = nutrient, y = total_wet_weight_g, fill = nutrient)) +
  geom_boxplot() +
  facet_wrap(~ light) +  
  scale_fill_npg(alpha = 0.5) +
  labs(y = "TotalBiomass (g)") +
  theme_pubr() +
  theme(legend.position = "top",
        axis.text.x = element_blank(),
    axis.title.x = element_blank())

#total biomass
plot7<- ggplot(metric_all, aes(x = nutrient, y = tissue_wet_weight_g, fill = nutrient)) +
  geom_boxplot() +
  facet_wrap(~ light) +  
  scale_fill_npg(alpha = 0.5) +
  labs(y = "TissueBiomass (g)") +
  theme_pubr() +
  theme(legend.position = "top",
        axis.text.x = element_blank(),
    axis.title.x = element_blank())

#cell density
plot8<- ggplot(metric_all, aes(x = nutrient, y = cell_count, fill = nutrient)) +
  geom_boxplot()+
  facet_wrap(~ light) +  
  scale_fill_npg(alpha = 0.5) +
  labs(y = "Symbiont cell density (10^7 cell g-1)") +
  theme_pubr() +
  theme(legend.position = "top",
        axis.text.x = element_blank(),
    axis.title.x = element_blank())

#chl-a
plot9<- ggplot(metric_all, aes(x = nutrient, y = chl_a, fill = nutrient)) +
  geom_boxplot() +
  facet_wrap(~ light) +  
  scale_fill_npg(alpha = 0.5) +
  labs(y = "Chl-a (ug g-1)") +
  theme_pubr() + 
  theme(legend.position = "top",
        axis.text.x = element_blank(),
    axis.title.x = element_blank())

fig6 <- (plot1+plot2+plot3+plot4+plot5+plot6+plot7+plot8+plot9) +
  plot_layout(ncol=4, guides = "collect") & 
  theme(legend.position = "top")
```
## Normality and homogeneity test
```{r}
#
load(file="output/growth_metric_all.RData")

#to check if data distribution is normal
options(digits=2)
metric_all %>% normality () %>% arrange(desc(p_value)) 
metric_all %>% normality () %>% filter(p_value >= 0.05) 

metric_all %>% plot_normality () #to visualize data distribution

find_skewness(metric_all, value=TRUE)#check for skewness value
find_skewness(metric_all, index=FALSE)#check for which variables have heavy skewness

#for leveneTest, use lapply across measurements
#combine light and nutrient level into single factor for grouping
metric_all <- metric_all %>%
  unite("trtmnt", light:nutrient, remove=FALSE)

#name trtmnt as factor
metric_all$trtmnt <- as.factor(metric_all$trtmnt)

#define function to perform Levene's test
levene_test_function <- function(variable) {
  leveneTest(variable ~ trtmnt, data = metric_all)
}

# List of variables to test
variables_to_test <- c("shell_length", "shell_width", "shell_height", 
                       "total_wet_weight_g", "tissue_wet_weight_g",
                       "shell_dry_weight_g", "strength_N", "cell_tissue_g", "chl_a_g")

# Apply the levene_test_function to each variable
results <- lapply(variables_to_test, function(var) {
  levene_test_function(metric_all[[var]])
})

# Name the list elements for easier interpretation
names(results) <- variables_to_test
```

### Data transformation
Data transformation of log10 or sqrt to check if data become normal
```{r}
#log10 transformation
metric_all_trans <- metric_all %>%
  mutate(across(shell_length:chl_a, ~ log1p(.)))

#check normality of log10 transformed data
metric_all_trans %>% normality () %>% filter(p_value >=0.05)
#adj_height, adj_strength, cell_count

#sqrt transformation
metric_all_trans2 <- metric_all %>%
  mutate(across(shell_length:chl_a, ~ sqrt(.)))

#check normality of log10 transformed data
metric_all_trans2 %>% normality () %>% filter(p_value >=0.05)

#log10 transformation
metric_all <- metric_all %>%
  mutate(log_cell_count = log10(cell_count))

#describe the cell count and chl-a
metric_all %>%
  group_by(light, nutrient)%>%
  describe() %>%
  filter(described_variables == "cell_count")
```
