---
title: "Part4-GLMM for growth performance"
author: "LeeLK"
date: "2025-04-08"
output: html_document
---
GLMM analysis for growth performance (SL growth, Fv/Fm, SGR)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Normality test
```{r}
library(dlookr)

#load data
load(file="output/growth_SL_long.RData")
load(file="output/pam.RData")
load(file="output/growth_rates.RData")

#check growth SL
growth_SL_long %>% plot_normality(SL)

#try transform with sqrt and log
growth_trans <- growth_SL_long %>%
  mutate(sqrt_SL= sqrt(SL)) %>%
  mutate(log_SL = log(SL))

#check normality
growth_trans %>% normality(log_SL, sqrt_SL)
#non-normal, proceed with GLMM

#check pam
pam %>% plot_normality(Y2)

#try transform with sqrt and log
pam_trans <- pam %>%
  mutate(sqrt_y2= sqrt(Y2)) %>%
  mutate(log_y2 = log(Y2))

#check normality
pam_trans %>% normality(log_y2, sqrt_y2)
#non-normal, proceed with GLMM

#check normality
growth_trans %>% normality(log_SL, sqrt_SL)
#non-normal, proceed with GLMM

#check sgr_t
growth_rate_changes_long %>% plot_normality(sgr_t)

#try transform with sqrt and log
sgr_trans <- growth_rate_changes_long %>%
  mutate(sqrt_sgr= sqrt(sgr_t)) %>%
  mutate(log_sgr = log(sgr_t))

#check normality
sgr_trans %>% normality(sqrt_sgr, log_sgr)
#non-normal, proceed with GLMM
```

## GLMM for variable with time
Light(PAR), nutrient and time (day) as fixed effect with interaction term. Tank as random effect to account for repeated measures across the 18 tanks

```{r}
#Package used
library(tidyverse)
library(lme4)
library(emmeans)
library(gtsummary)
library(sjPlot)
library(car)
library(multcomp)
library(ggpubr)
library(ggsci)
library(dlookr)#for quick normality

#since SL is non-normal, Generalized linear mixed model will be used
#treat time as polynomial with 2 degree consider, in this case, with both linear and quadratic effects, a more complex, curvilinear relationship between time and shell length growth
#because it is expect that the effect of time on SL growth isnt strictly linear but may increase or decrease at varying rate over time (likely decrease)

#load growth SL data
load(file="output/growth_SL_long.RData")

set.seed(1234)
#backward-stepwise model simplification
#Start with full model
glmm_SL <- glmer(SL ~ light*nutrient*poly(time_days,2) + (1|tank), 
                     family = Gamma(link ="log"),
                    data =growth_SL_long,
                    control = glmerControl(optimizer = "bobyqa", 
                                           optCtrl = list(maxfun = 1e5)))
#no 3-way interaction
glmm_SL_no_3way <- glmer(SL ~ (light + nutrient + poly(time_days, 2))^2 
                         + (1 | tank),
                         family = Gamma(link = "log"),
                         data = growth_SL_long,
                         control = glmerControl(optimizer = "bobyqa", 
                                                optCtrl = list(maxfun = 1e5)))
#only light and nutrient
glmm_SL_lightnutrient <- glmer(SL ~ light*nutrient + (1 | tank),
                       family = Gamma(link = "log"),
                       data = growth_SL_long,
                       control = glmerControl(optimizer = "bobyqa", 
                                              optCtrl = list(maxfun = 1e5)))

#only light and time
glmm_SL_lighttime <- glmer(SL ~ light*poly(time_days,2) + (1 | tank),
                       family = Gamma(link = "log"),
                       data = growth_SL_long,
                       control = glmerControl(optimizer = "bobyqa", 
                                              optCtrl = list(maxfun = 1e5)))
#only nutrient and time
glmm_SL_nutrienttime <- glmer(SL ~ nutrient*poly(time_days,2) + (1 | tank),
                       family = Gamma(link = "log"),
                       data = growth_SL_long,
                       control = glmerControl(optimizer = "bobyqa", 
                                              optCtrl = list(maxfun = 1e5)))

#Run AIC for model comparison
AIC_r1 <- AIC(glmm_SL, glmm_SL_no_3way,glmm_SL_lightnutrient, 
              glmm_SL_lighttime, glmm_SL_nutrienttime)
#glmm_SL, with 3-way interaction with lowest AIC, hence this model is selected

#to compare AIC model with delta AIC
models <- list(full= glmm_SL, no_3way = glmm_SL_no_3way,
               light_nutrient = glmm_SL_lightnutrient, 
              light_time = glmm_SL_lighttime, 
              nutrient_time = glmm_SL_nutrienttime)
aic_values <- sapply(models, AIC)
delta_AIC <- aic_values - min(aic_values)
SLmodel_AIC_comp <- data.frame(Model = names(models), AIC = aic_values, delta_AIC = delta_AIC)

#write csv as output
write.csv(AIC_r1, file="temp_output/AIC_sl.csv")

#summarise selected model with gtsummary table format
glmm_SL_summary <- tbl_regression(glmm_SL)

#use sjPLot to extract the result on random effect
tab_model(glmm_SL)

#use x2 statistic to assess significance of fixed effects either
#likelihood ratio tests (LRT) to compared reduced model and full model
#Wald tests for each fixed effects

#Type 3 ANOVA for unbalanced design
Anova(glmm_SL, type=3)

#post-hoc test regardless of anova results, as we are interest in specific time points

#all time points, specifically, at 30, 60, 90, 120, 150
emms_SL<- emmeans(glmm_SL, pairwise ~ light*nutrient*poly(time_days,2), at = list(time_days = c(30,60,90,120,150)), type ="response", adjust = "bonferroni")
#alot of significant due to the fact that time is a significant effect on growth

summary(emms_SL)

#extract by time point, day 30
emms_SL_d30<- emmeans(glmm_SL, pairwise ~ light*nutrient*poly(time_days,2), at = list(time_days = 30), type ="response", adjust = "bonferroni")

emms_SL30 <- cld(emms_SL_d30, alpha=0.05, Letters=letters, adjust="bonferroni")

emms_SL_d30df <- as.data.frame(summary(emms_SL_d30))

write.csv(emms_SL_d30df, file="output/emms_SL_d30df.csv")

#extract by time point, day 60
emms_SL_d60<- emmeans(glmm_SL, pairwise ~ light*nutrient*poly(time_days,2), at = list(time_days = 60), type ="response", adjust = "bonferroni")

emms_SL60 <- cld(emms_SL_d60, alpha=0.05, Letters=letters, adjust="bonferroni")

emms_SL_d60df <- as.data.frame(summary(emms_SL_d60))

write.csv(emms_SL_d60df, file="output/emms_SL_d60df.csv")

#extract by time point, day 90
emms_SL_d90<- emmeans(glmm_SL, pairwise ~ light*nutrient*poly(time_days,2), at = list(time_days = 90), type ="response", adjust = "bonferroni")

emms_SL90 <- cld(emms_SL_d90, alpha=0.05, Letters=letters, adjust="bonferroni")

emms_SL_d90df <- as.data.frame(summary(emms_SL_d90))

write.csv(emms_SL_d90df, file="output/emms_SL_d90df.csv")

#extract by time point, first day 120
emms_SL_d120<- emmeans(glmm_SL, pairwise ~ light*nutrient*poly(time_days,2), at = list(time_days = 120), type ="response", adjust = "bonferroni")

emms_SL120 <- cld(emms_SL_d120, alpha=0.05, Letters=letters, adjust="bonferroni")

emms_SL_d120df <- as.data.frame(summary(emms_SL_d120))

write.csv(emms_SL_d120df, file="output/emms_SL_d120df.csv")

#extract by time point, day 150
emms_SL_d150<- emmeans(glmm_SL, pairwise ~ light*nutrient*poly(time_days,2), at = list(time_days = 150), type ="response", adjust = "bonferroni")

emms_SL150 <- cld(emms_SL_d150, alpha=0.05, Letters=letters, adjust="bonferroni")

emms_SL_d150df <- as.data.frame(summary(emms_SL_d150))

write.csv(emms_SL_d150df, file="output/emms_SL_d150df.csv")

#bind all data
emms_SL30_150 <- rbind(emms_SL30, emms_SL60, emms_SL90, 
                       emms_SL120, emms_SL150)

emms_SL30_150 <- emms_SL30_150 %>%
  unite("trtmnt", light, nutrient) %>%
  mutate_at(vars(trtmnt, time_days), factor)

emms_SL30_150$time_days <- factor(emms_SL30_150$time_days, levels = c(150, 120, 90, 60, 30))

save(emms_SL30_150, file = "output/emms_SLplot.RData")

#Fig2b: Pointrange plot
emms_plot1 <-ggplot(emms_SL30_150, aes(x=trtmnt, y=response))+
  geom_pointrange(aes(ymin = asymp.LCL, ymax = asymp.UCL, color=trtmnt)) +
  facet_grid(rows = vars(time_days))+
  scale_color_frontiers()+
  geom_text(aes(label = .group), vjust = -0.5, hjust = 0.5, size = 10/.pt) + 
  coord_flip()+
  labs(x= "Light:Nutrient:Time", y = "Estimated ShLength (mm)")+
  theme_pubr() +
  theme(axis.text.y = element_blank(),
        legend.position = c(0.8,0.2))
```

## Assemble Fig2

```{r}
library(patchwork)
#plot1 from part 3
#assembly plot for fig3
fig2 <- (plot1|emms_plot1) + plot_annotation(tag_levels = "a")
```

## GLMM for specifc growth rate changes over time

```{r}
#load
load(file="output/growth_rates.RData")
set.seed(1234)

#since sgr also not conform to normality, glmm was used
#similar to SL growth, time was considered as polynomial 
#full model
glmm_sgr <- glmer(sgr_t ~ light*nutrient*poly(time_days,2) 
                  + (1|tank), family = Gamma(link ="log"),
                    data =growth_rate_changes_long, 
                  control = glmerControl(optimizer = "bobyqa", 
                                         optCtrl = list(maxfun = 1e5)))
#model comparison
glmm_sgr_no_3way <- glmer(sgr_t ~ (light + nutrient + poly(time_days, 2))^2 
                         + (1 | tank),
                         family = Gamma(link = "log"),
                         data = growth_rate_changes_long,
                         control = glmerControl(optimizer = "bobyqa", 
                                                optCtrl = list(maxfun = 1e5)))
#no 3 way
glmm_sgr_lightnutrient <- glmer(sgr_t ~ light*nutrient + (1 | tank),
                       family = Gamma(link = "log"),
                       data = growth_rate_changes_long,
                       control = glmerControl(optimizer = "bobyqa", 
                                              optCtrl = list(maxfun = 1e5)))
#light and time
glmm_sgr_lighttime <- glmer(sgr_t ~ light*poly(time_days,2) + (1 | tank),
                       family = Gamma(link = "log"),
                       data = growth_rate_changes_long,
                       control = glmerControl(optimizer = "bobyqa", 
                                              optCtrl = list(maxfun = 1e5)))
#nutrient and time
glmm_sgr_nutrienttime <- glmer(sgr_t ~ nutrient*poly(time_days,2) + (1 | tank),
                       family = Gamma(link = "log"),
                       data = growth_rate_changes_long,
                       control = glmerControl(optimizer = "bobyqa", 
                                              optCtrl = list(maxfun = 1e5)))
#Compute AIC
AIC_sgr <- AIC(glmm_sgr, glmm_sgr_no_3way,glmm_sgr_lightnutrient, 
              glmm_sgr_lighttime, glmm_sgr_nutrienttime)

#glmm_sgr_nutrienttime was selected with lowest AIC

#Compare AIC model with delta AIC
models <- list(full= glmm_sgr, no_3way = glmm_sgr_no_3way,
               light_nutrient = glmm_sgr_lightnutrient, 
              light_time = glmm_sgr_lighttime, 
              nutrient_time = glmm_sgr_nutrienttime)
aic_values <- sapply(models, AIC)
delta_AIC <- aic_values - min(aic_values)
sgrmodel_AIC_comp <- data.frame(Model = names(models), AIC = aic_values, delta_AIC = delta_AIC)

#write csv as output
write.csv(sgrmodel_AIC_comp, file="output/sgrmodel_AIC_comp_SL.csv")

# Summary of the model
summary(glmm_sgr_nutrienttime)

#summarise with gtsummary table format
glmm_sgr_summary <- tbl_regression(glmm_sgr_nutrienttime)

#use sjPLot to extract the result on random effect
tab_model(glmm_sgr_nutrienttime)

#use x2 statistic to assess significance of fixed effects either
#likelihood ratio tests (LRT) to compared reduced model and full model
#Wald tests for each fixed effects

#ANOVA for nutrient-time model
Anova(glmm_sgr_nutrienttime, type =3)

#all time points, specifically, at 30, 60, 120, 180
emms_sgr_all<- emmeans(glmm_sgr_nutrienttime, pairwise ~ nutrient*poly(time_days,2), at = list(time_days = c(30,60,90,120,150)), type ="response", adjust = "bonferroni")
#alot of significant due to the fact that time is a significant effect on growth

summary(emms_sgr_all)

#extract by time point, first day 30
emms_sgr_d30<- emmeans(glmm_sgr_nutrienttime, pairwise ~ nutrient*poly(time_days,2), at = list(time_days = 30), type ="response", adjust = "bonferroni")

emms_sgr30 <- cld(emms_sgr_d30, alpha=0.05, Letters=letters, adjust="bonferroni")

#extract by time point, first day 60
emms_sgr_d60<- emmeans(glmm_sgr_nutrienttime, pairwise ~ nutrient*poly(time_days,2), at = list(time_days = 60), type ="response", adjust = "bonferroni")

emms_sgr60 <- cld(emms_sgr_d60, alpha=0.05, Letters=letters, adjust="bonferroni")

#extract by time point, first day 90
emms_sgr_d90<- emmeans(glmm_sgr_nutrienttime, pairwise ~ nutrient*poly(time_days,2), at = list(time_days = 90), type ="response", adjust = "bonferroni")

emms_sgr90 <- cld(emms_sgr_d90, alpha=0.05, Letters=letters, adjust="bonferroni")

#extract by time point, first day 120
emms_sgr_d120<- emmeans(glmm_sgr_nutrienttime, pairwise ~ nutrient*poly(time_days,2), at = list(time_days = 120), type ="response", adjust = "bonferroni")

emms_sgr120 <- cld(emms_sgr_d120, alpha=0.05, Letters=letters, adjust="bonferroni")

#extract by time point, first day 150
emms_sgr_d150<- emmeans(glmm_sgr_nutrienttime, pairwise ~ nutrient*poly(time_days,2), at = list(time_days = 150), type ="response", adjust = "bonferroni")

emms_sgr150 <- cld(emms_sgr_d150, alpha=0.05, Letters=letters, adjust="bonferroni")

#bind all cld data
emms_sgr_all<- rbind(emms_sgr30, emms_sgr60, emms_sgr90, 
                       emms_sgr120, emms_sgr150)

emms_sgr_all <- emms_sgr_all %>%
  unite("trtmnt", nutrient) %>%
  mutate_at(vars(trtmnt, time_days), factor)

emms_sgr_all$time_days <- factor(emms_sgr_all$time_days, levels = c(150, 120, 90, 60, 30))

save(emms_sgr_all, file="output/emms_sgrPlot.RData")
#
emms_plot2 <-ggplot(emms_sgr_all, aes(x=trtmnt, y=response))+
  geom_pointrange(aes(ymin = asymp.LCL, ymax = asymp.UCL, color=trtmnt)) +
  facet_grid(rows = vars(time_days))+
  scale_color_npg() +
  geom_text(aes(label = .group), vjust = -0.5, hjust = 0.5, size = 10/.pt) + 
  coord_flip()+
  labs(x= "Nutrient:Time", y = "Estimated SGR (day-1)")+
  theme_pubr() +
  theme(axis.text.y = element_blank(),
        legend.position = c(0.8,0.8))
```

## Assemble Fig3

```{r}
library(patchwork)
#assembly plot for fig3
fig3 <- (plot3|emms_plot2) + plot_annotation(tag_levels = "a")
```

## GLMM for photosynthetic efficiency

```{r}
load(file="output/pam.RData")

#full model
glmm_pam <- glmer(Y2 ~ light*nutrient*poly(time_days,2) + (1|tank), 
                     family = Gamma(link ="log"),
                    data =pam, 
                  control = glmerControl(optimizer = "bobyqa", 
                                         optCtrl = list(maxfun = 1e5)))

#model comparison
glmm_pam_no_3way <- glmer(Y2 ~ (light + nutrient + poly(time_days, 2))^2 
                         + (1 | tank),
                         family = Gamma(link = "log"),
                         data = pam,
                         control = glmerControl(optimizer = "bobyqa", 
                                                optCtrl = list(maxfun = 1e5)))
#light and nutrient
glmm_pam_lightnutrient <- glmer(Y2 ~ light*nutrient + (1 | tank),
                       family = Gamma(link = "log"),
                       data = pam,
                       control = glmerControl(optimizer = "bobyqa", 
                                              optCtrl = list(maxfun = 1e5)))
#light and time
glmm_pam_lighttime <- glmer(Y2 ~ light*poly(time_days,2) + (1 | tank),
                       family = Gamma(link = "log"),
                       data = pam,
                       control = glmerControl(optimizer = "bobyqa", 
                                              optCtrl = list(maxfun = 1e5)))
#nutrient and time
glmm_pam_nutrienttime <- glmer(Y2 ~ nutrient*poly(time_days,2) + (1 | tank),
                       family = Gamma(link = "log"),
                       data = pam,
                       control = glmerControl(optimizer = "bobyqa", 
                                              optCtrl = list(maxfun = 1e5)))
#AIC run
AIC_pam <- AIC(glmm_pam, glmm_pam_no_3way,glmm_pam_lightnutrient, 
              glmm_pam_lighttime, glmm_pam_nutrienttime)
#full model, with 3-way interaction with lowest AIC, hence this model is selected

#to formally compare AIC model with delta AIC
models <- list(full= glmm_sgr, no_3way = glmm_sgr_no_3way,
               light_nutrient = glmm_sgr_lightnutrient, 
              light_time = glmm_sgr_lighttime, 
              nutrient_time = glmm_sgr_nutrienttime)
aic_values <- sapply(models, AIC)
delta_AIC <- aic_values - min(aic_values)
sgrmodel_AIC_comp <- data.frame(Model = names(models), AIC = aic_values, delta_AIC = delta_AIC)

#write csv as output
write.csv(sgrmodel_AIC_comp, file="output/sgrmodel_AIC_comp_SL.csv")

#summarise with gtsummary table format
glmm_pam_summary <- tbl_regression(glmm_pam)

#use sjPLot to extract the result on random effect
tab_model(glmm_pam)

#use x2 statistic to assess significance of fixed effects either
#likelihood ratio tests (LRT) to compared reduced model and full model
#Wald tests for each fixed effects
Anova(glmm_pam, type = 3)
#

#extract by time point, first day 30
emms_pam_d30<- emmeans(glmm_pam, pairwise ~ light*nutrient*poly(time_days,2), at = list(time_days = 30), type ="response", adjust = "bonferroni")

emms_pam30 <- cld(emms_pam_d30, alpha=0.05, Letters=letters, adjust="bonferroni")

#extract by time point, first day 60
emms_pam_d60<- emmeans(glmm_pam, pairwise ~ light*nutrient*poly(time_days,2), at = list(time_days = 60), type ="response", adjust = "bonferroni")

emms_pam60 <- cld(emms_pam_d60, alpha=0.05, Letters=letters, adjust="bonferroni")

#extract by time point, first day 90
emms_pam_d90<- emmeans(glmm_pam, pairwise ~ light*nutrient*poly(time_days,2), at = list(time_days = 90), type ="response", adjust = "bonferroni")

emms_pam90 <- cld(emms_pam_d90, alpha=0.05, Letters=letters, adjust="bonferroni")

#extract by time point, first day 120
emms_pam_d120<- emmeans(glmm_pam, pairwise ~ light*nutrient*poly(time_days,2), at = list(time_days = 120), type ="response", adjust = "bonferroni")

emms_pam120 <- cld(emms_pam_d120, alpha=0.05, Letters=letters, adjust="bonferroni")

#extract by time point, first day 30
emms_pam_d150<- emmeans(glmm_pam, pairwise ~ light*nutrient*poly(time_days,2), at = list(time_days = 150), type ="response", adjust = "bonferroni")

emms_pam150 <- cld(emms_pam_d150, alpha=0.05, Letters=letters, adjust="bonferroni")

#rbind all the dataset
emms_pam_all <- rbind(emms_pam30, emms_pam60, emms_pam90, 
                       emms_pam120, emms_pam150)

emms_pam_all <- emms_pam_all %>%
  unite("trtmnt", light, nutrient) %>%
  mutate_at(vars(trtmnt, time_days), factor)

emms_pam_all$time_days <- factor(emms_pam_all$time_days, levels = c(150, 120, 90, 60, 30))

save(emms_pam_all, file ="output/emms_pamPlot.RData")

#
emms_plot3 <-ggplot(emms_pam_all, aes(x=trtmnt, y=response))+
  geom_pointrange(aes(ymin = asymp.LCL, ymax = asymp.UCL, color=trtmnt)) +
  facet_grid(rows = vars(time_days))+
  scale_color_frontiers()+
  geom_text(aes(label = .group), vjust = -0.5, hjust = 0.5, size = 10/.pt) + 
  coord_flip()+
  labs(x= "Light:Nutrient:Time", y = "Estimated Fv/Fm")+
  theme_pubr() +
  theme(axis.text.y = element_blank(),
        legend.position = c(0.8,0.8))
```

## Assemble Fig5

```{r}
library(patchwork)
#assembly plot for fig3
fig5 <- (plot2|emms_plot3) + plot_annotation(tag_levels = "a")
```
